SPANG: constructing a controlled set of queries with the FAIR principle

### 2020-04-06 paper.tex
Updates from Overleaf

\title{SPANG: a controlled query set with the FAIR principle}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
Here we overview the structure of this paper.

@@

%%%%%%%%%%%%%%%%%%%%%%%%

\section{SPANG}
SPANG is originally developed for "SARQLing" the RDF dataset from command line. Here, we re-implemented the SPANG using Node.js and now the functionality is available through Web browser.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG library}

@@

      ]
    },
    {
      "name": "gene_disease",
      "title": "Get disease associated with NCBI gene ID",
      "uri": "http://localhost:7070/api/disgenet/gene_disease",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [
        {
          "name": "arg1",
          "default": "675"
        }
      ]
    },
    {
      "name": "get_gda",
      "title": "Get specific type of gene-disease association",
      "uri": "http://localhost:7070/api/disgenet/get_gda",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [
        {
          "name": "arg1",
          "default": "SIO_001120"
        },
        {
          "name": "arg2",
          "default": "10"
        }
      ]
    },
    {
      "name": "list_gda_sources",
      "title": "List sources of gene-disease association",
      "uri": "http://localhost:7070/api/disgenet/list_gda_sources",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [

      ]
    },
    {
      "name": "list_gene_disease",
      "title": "List gene disease associations",
      "uri": "http://localhost:7070/api/disgenet/list_gene_disease",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [
        {
          "name": "arg1",
          "default": "10"
        }
      ]
    },
    {
      "name": "regex_disease",
      "title": "Search the NCI Thesaurus for specific diseases",
      "uri": "http://localhost:7070/api/disgenet/regex_disease",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [
        {
          "name": "arg1",
          "default": "breast"
        }
      ]
    },
    {
      "name": "regex_gda_disease",
      "title": "Regex seach for specific diseases that have gene-disease association information",
      "uri": "http://localhost:7070/api/disgenet/regex_gda_disease",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [
        {
          "name": "arg1",
          "default": "alzheimer"
        }
      ]
    },
    {
      "name": "uniprot2disease",
      "title": "Get diseases associated with given UniProt IDs",
      "uri": "http://localhost:7070/api/disgenet/uniprot2disease",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [

      ]
    },
    {
      "name": "uniprot2disease_curated",
      "title": "Get diseases associated with given UniProt IDs",
      "uri": "http://localhost:7070/api/disgenet/uniprot2disease_curated",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [

      ]
    },
    {
      "name": "uniprot2disease_curated_none",
      "title": "Get diseases associated with given UniProt IDs",
      "uri": "http://localhost:7070/api/disgenet/uniprot2disease_curated_none",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [

      ]
    },
    {
      "name": "uniprot2disease_none",
      "title": "Get diseases associated with given UniProt IDs",
      "uri": "http://localhost:7070/api/disgenet/uniprot2disease_none",
      "endpoint": "http://rdf.disgenet.org/sparql/",
      "param": [

      ]
    }
  ],
  "general": [
    {
      "name": "find_term_uri",
      "title": "Find classes or properties (regex search for URIs or labels)",
      "uri": "http://localhost:7070/api/general/find_term_uri",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [
        {
          "name": "arg1",
          "default": "homolo"
        }
      ]
    },
    {
      "name": "graphs_bug",
      "title": "SELECT DISTINCT ?graph, instead of GROUP BY ?graph, returns less results (Virtuoso 7.1.0)",
      "uri": "http://localhost:7070/api/general/graphs_bug",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "label",
      "title": "",
      "uri": "http://localhost:7070/api/general/label",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "link_select",
      "title": "Links from a specific instance",
      "uri": "http://localhost:7070/api/general/link_select",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "regex_uri",
      "title": "Regex search for URIs (subjects of triples)",
      "uri": "http://localhost:7070/api/general/regex_uri",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "search_for_class",
      "title": "Free text search in Virtuoso",
      "uri": "http://localhost:7070/api/general/search_for_class",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "search_label_for_class",
      "title": "Free text search in Virtuoso",
      "uri": "http://localhost:7070/api/general/search_label_for_class",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    },
    {
      "name": "search_label_not_class",
      "title": "Free text search in Virtuoso (count instances of each class)",
      "uri": "http://localhost:7070/api/general/search_label_not_class",
      "endpoint": "http://dbpedia.org/sparql",
      "param": [

      ]
    }
  ]
}

@@

%%%%%%%%%%%%%%%%%%%%
\section{Use cases}

@@

\noindent where the first argument is the target SPARQL endpoint and the following arguments are SPARQL shortcuts (the SPARQL endpoint can be specified in a URL or nickname for simplicity). 
\texttt{uniprot} is a nickname for the UniProt SPARQL endpoint \citep{UniProt} and \texttt{tax:} is a prefix for URIs of NCBI taxonomy IDs. 
This example command line searches the target database for statements that have taxonomy ID 511145 ({\it Escherichia coli} strain K-12 MG1655) as a subject. Using the \texttt{-a} option transforms the URIs in the
 search result into abbreviated forms using predefined prefix declarations. For example,

 \texttt{\textless http://www.w3.org/2000/01/rdf-schema\#subClassOf\textgreater}

\noindent is transformed into \texttt{rdfs:subClassOf
}.
The result is output to the standard output in the form of tab-separated values by default, as it is suitable for processing by line-oriented Unix programs.
In addition, combination of subject, predicate, and object is possible according to the following.

>>>>

\title{SPANG: constructing a controlled set of queries with the FAIR principle}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{The FAIR principle}

Not only the RDF dataset, but also the SPARQL queries are important resource to be shared on the Semantic Web platform. Thus, we applied the FAIR principle which have been applied to the RDF datasets, into SPARQL queries.

Here we overview the preliminaries for the FAIR principle.
It shoulbe noted that the FIAR principle can be applied not only to the dataset resources but also to other type of resources including software tools.

Here, we provide the FAIR principle in the context of SPARQL queries.

\subsection{Findable}
\begin{itemize}
    \item Assign global identifier
    \item Rich metadata
    \item Include the identifier of the data they describe
    \item Registered and indexed
\end{itemize}

\subsection{Accessible}
\begin{itemize}
    \item Retrievable by identifier
    \item Metadata are accessible
\end{itemize}

\subsection{Interoperable}
\begin{itemize}
    \item Use a formal, accessible, shared, and broadly applicable language for knowledge representation.
    \item Use vocabularies that follow FAIR principles.
    \item Include qualified references to others
\end{itemize}

\subsection{Reusable}
\begin{itemize}
    \item Richly described with a plurality of accurate and relevant attributes
    \begin{itemize}
        \item released with a clear and accessible data usage license
        \item associated with detailed provenance
        \item meet domain-relevant community standards
    \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG overview}
Here we overview the structure of this paper.

@@

\subsection{SPANG portal}
SPANG is originally developed for "SARQLing" the RDF dataset from command line. Here, we re-implemented the SPANG using Node.js and now the functionality is available through Web browser.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG specifications}

@@

      ]
    },
    ...
  ]
}

@@

%%%%%%%%%%%%%%%%%%%%

@@

\noindent where the first argument is the target SPARQL endpoint and the following arguments are SPARQL shortcuts (the SPARQL endpoint can be specified in a URL or nickname for simplicity). 
\texttt{uniprot} is a nickname for the UniProt SPARQL endpoint \citep{UniProt} and \texttt{tax:} is a prefix for URIs of NCBI taxonomy IDs. 
This example command line searches the target database for statements that have taxonomy ID 511145 ({\it Escherichia coli} strain K-12 MG1655) as a subject. Using the \texttt{-a} option transforms the URIs in the
 search result into abbreviated forms using predefined prefix declarations. For example,\\
\texttt{\textless http://www.w3.org/2000/01/rdf-schema\#subClassOf\textgreater} 
is transformed into \texttt{rdfs:subClassOf
}.
The result is output to the standard output in the form of tab-separated values by default, as it is suitable for processing by line-oriented Unix programs.
In addition, combination of subject, predicate, and object is possible according to the following.
